---
title: "TidyTemplate"
date: 2020-10-26
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, cache = TRUE,
                      warning = FALSE, message = FALSE,
                      dpi = 180, fig.height = 5, fig.width = 8)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2020-09-15")
kids_raw <- tt$kids

kids_raw %>% 
    saveRDS("kids_raw.RDS")

kids_raw <- read_rds("kids_raw.RDS")

variable_choice <- kids %>% 
    distinct(variable) %>% 
    bind_cols(var_name = c("elementary and secondary education",
                           "higher education",
                           "education subsidies",
                           "education special services",
                           "Pell Grants",
                           "Federal Head Start",
                           "TANF",
                           "other cash assistance",
                           "SNAP benefit",
                           "Social Security",
                           "SSI payments",
                           "EITC",
                           "Child Tax Credit",
                           "Additional Child Tax credit",
                           "State_EITC",
                           "Unemployment Compensation",
                           "Workers Compensation",
                           "Medicaid_CHIP",
                           "public health",
                           "Other health",
                           "housing and community development",
                           "libraries",
                           "parks and recreation"))

# Removing na observations from the raw data
kids <- kids_raw %>% 
    drop_na() %>% 
    inner_join(variable_choice)

# Saving the final clean data for future use
kids %>% 
    saveRDS("kids.RDS")
```

Now, load the clean data
```{r}
kids <- read_rds("kids.RDS")
kids

```

Govt Speding on Elementary Education and Higher Education
```{r}
kids_ele_high <- kids %>% 
    filter(variable %in% c("PK12ed", "highered"))

kids_ele_high %>% 
    group_by(variable) %>% 
    summarise(total = sum(inf_adj_perchild))
    
```

* Govt has spent ~5 times more on primary/elementary education than higher education in total in last 20 Years
```{r}
kids_ele_high %>% 
    group_by(variable, year) %>% 
    summarise(total = sum(inf_adj_perchild)) %>% 
    ggplot(aes(year, total)) +
    geom_line(aes(color = variable))

```

```{r}
kids %>% 
    filter(variable == "unemp") %>% 
    group_by(year) %>% 
    summarise(total = sum(inf_adj_perchild)) %>% 
    ggplot(aes(year, total)) +
    geom_line(size =1, color = "gray60")
```

Expenditure on public health and other health
```{r}
kids_health <- kids %>% 
    filter(variable %in% c("pubhealth", "other_health"))

kids_health %>% 
    group_by(variable, year) %>% 
    summarise(total = sum(inf_adj_perchild)) %>% 
    ggplot(aes(year, total)) +
    geom_line(aes(color = variable))
        
```

Other health has a significant expenditure. It also has a constant linear trend which means that expenditure is increasing continuously over years, public health also shows increase bit the slope is almost flat.

Public Expenditure on libraries
```{r}
kids %>% 
    filter(variable == "lib") %>% 
    group_by(year) %>% 
    summarise(total = sum(inf_adj_perchild)) %>% 
    ggplot(aes(year, total)) +
    geom_line(size =1, color = "gray60")
```

Public Expenditure on parks and recreational spaces
```{r}
kids %>% 
    filter(variable == "parkrec") %>% 
    group_by(year) %>% 
    summarise(total = sum(inf_adj_perchild)) %>% 
    ggplot(aes(year, total)) +
    geom_line(size =1, color = "gray60") +
    labs(title = "Gov Spendings on Parks and Recreational Spaces",
         x = "",
         y = "Total Amount in $ per child")
```

Lets create a shiny app to see these trends in more interactive form
Plot for shiny app
```{r}
kids %>% 
    filter(variable == "parkrec",
           state == "Alabama") %>%
    group_by(year) %>% 
    summarise(total = sum(inf_adj_perchild)) %>% 
    ggplot(aes(year, total)) +
    geom_line(size =1, color = "gray60") +
    labs(title = "Gov Spendings on Parks and Recreational Spaces",
         x = "",
         y = "Total Amount in $ per child")
```

